Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-24 paper/main.tex
qualitative evaluation examples; graph

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={baseline, Type, Attribute, Value, Key},
    xtick=data,
    ymin=0,
    ylabel=Model Accuracy,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=score, col sep=comma]{defect-categories-2.csv};
\end{axis}
\end{tikzpicture}
\caption{Accuracy when the model is trained and tested on only the data exhibiting
the error on the x-axis.
\fixme{Spin this for generality. Indicate that our model gives high-quality
fault localization regardless of defect type. To put it another way, we
can view this as a sensitivity analysis: our algorithm is not sensitive to
defect type (that's good!).}
}
\label{fig:defect-categories}
\end{figure}

@@

Pick out one example where we did poorly. Walk through why it is ``out of
scope'', in some sense: we would need features that capture XYZ to handle
it correctly, but those are either expensive or undecideable. Hint at
future work.

Pick out a few examples where we did well and the baseline did poorly. Walk
through how interesting and indicative they are and explain why we did well.

>>>>

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={baseline, Type, Attribute, Value, Key, Index},
    xtick=data,
    ymin=0,
    ylabel=Model Accuracy,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=score, col sep=comma]{defect-categories-2.csv};
\end{axis}
\end{tikzpicture}
\caption{Accuracy when the model is trained and tested on only the data exhibiting
the error on the x-axis.
\fixme{Spin this for generality. Indicate that our model gives high-quality
fault localization regardless of defect type. To put it another way, we
can view this as a sensitivity analysis: our algorithm is not sensitive to
defect type (that's good!).}
}
\label{fig:defect-categories}
\end{figure}

@@

\fixme{
Pick out one example where we did poorly. Walk through why it is ``out of
scope'', in some sense: we would need features that capture XYZ to handle
it correctly, but those are either expensive or undecideable. Hint at
future work.
}

%%% Original for fig-win-example-1; index 19615 of v4
% def word_checker(food):
%     w_alphabet_ned = ['a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j',
%                 'k', 'l', 'm', 'n', 'o', 'p', 'q', 'r', 's', 't',
%                 'u', 'v', 'w', 'x', 'y', 'z', 'ë', 'ï', 'é', 'è', 'ö', 'ê',
%                 'ü', 'ó', 'ç', 'á', 'à', 'ä', 'û', 'î', 'í', 'ô', 'ú', 'ñ', 'â']
%
%     w_list_food = list(food)
%
%     for letters in w_list_food:
%         if letters not in w_alphabet_ned:
%             w_list_food = w_list_food.remove(letters)
%         food = ''.join(w_list_food)
%     print(food)
%
%
%
%
% word_checker('sl^* ( ) 0a')

\begin{figure}
\begin{lstlisting}
def devowel(word):
    word = 'foo'
    w_list = list(word)
    vowels = ['a', 'e', 'i', 'o', 'u']
    for letter in w_list:
        if letter not in vowels:
            w_list = w_list.remove(letter)
    word = ''.join(w_list)
print(devowel('foo'))
\end{lstlisting}
\caption{A program with a bug on line 7.}
\label{fig-win-example-one}
\end{figure}

%%% Original for fig-win-example-2; index 50219 of v4
% GeorgiaAreaCodes = [404, 770, 678, 706, 762, 229, 423, 912, 478]
% for i in range(0, len(GeorgiaAreaCodes) + 1):
%     print(GeorgiaAreaCodes[i])

\begin{figure}
\begin{lstlisting}
areaCodes = [800, 555]
for i in range(0, len(areaCodes) + 1):
    print(areaCodes[i])
\end{lstlisting}
\caption{A program with an off-by-one bug.}
\label{fig-win-example-two}
\end{figure}

We now go through a couple examples where our method does well. These examples
are adapted slightly for simplicity of
presentation and to protect the anonymity of the programmers, but they retain
the essential character of the originals from our dataset.

\paragraph{NoneType}
The function in \figref{win-example-one} attempts to remove all vowels from a
given word. However the
\lt{remove} method in line 7 actually modifies the list in place and returns
\lt{None}, so the correct version forgoes the assignment and just has
\lt{w_list.remove(letter)}. Python does not crash until line 8, where its
message is \lt{TypeError: can only join an iterable}. However, our decision
tree flags the correct statement, using as some of its rules that it is an
assignment statement whose second child has type \lt{NoneType}. This seems to
capture the intuitive rule that there is rarely a good reason to assign the
value \lt{None} to a variable. Note that these rules use both the syntactic
form of the statement (an assignment) and the dynamic type of one of its
children, so all our categories of features - syntactic, dynamic, and
contextual - were useful here.

\paragraph{Off-by-one bugs}
In \figref{win-example-two}, the programmer incorrectly adds one to the high
end of their range, causing an \lt{IndexError: list index out of range} on line
3 during the final iteration of the for loop. The correct expression to blame
is the addition \lt{len(areaCodes) + 1}. Some of the rules that our
decision tree uses to correctly localize this bug include that the error is an
index error, the type of the parent, and the fact that it's an addition
expression. Thus once again we see all three categories of features being used.
